---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
  )
```

# PALIN: Convert your PAs into INs

<!-- badges: start -->
<!-- badges: end -->

The goal of `palin` is to provide utilities for working with reverse correlation data. It can be used, for instance, to estimate internal noise in such tasks.

## Installation

You can install the development version of `palin` from GitHub with:

```{r, eval = FALSE}
install.packages("remotes")

remotes::install_github(
    repo = "https://github.com/neuro-team-femto/palin/r",
    dependencies = TRUE
    )
```

## Usage

Computing response bias and internal noise from the percentage of agreement.

```{r sdt-example, eval = TRUE}
library(palin)

# example data (70% agreement and no response bias)
df <- data.frame(prop_agree = 0.7, prop_int1 = 0.5, ntrials = 1e2)

# fitting the SDT model to these data
sdt_fitting(data = df, ntrials = 1e4, method = "bobyqa")
```

Fitting the drift diffusion model to the raw data (minus the last block).

```{r ddm-example, eval = TRUE, message = FALSE}
library(tidyverse)

# importing the data
data(self_produced_speech)
head(self_produced_speech)

# reshaping the data
df <- self_produced_speech |>
    # removing the last block
    filter(block < 7) |>
    # keeping only the relevant columns
    select(participant, choice = response, RT) |>
    mutate(choice = ifelse(test = choice == "stim1", yes = 1, no = 2) )

# splitting data by participant
df_ln <- df |> filter(participant == "LN")
df_mm <- df |> filter(participant == "MM")

# fitting the full DDM for LN and MM (pars are a, v, t0, w, sv)
ddm_fitting(rt = df_ln$RT, resp = df_ln$choice, method = "nlminb")$par
ddm_fitting(rt = df_mm$RT, resp = df_mm$choice, method = "nlminb")$par
```
